---
title: "Absenteeism-at-work"
subtitle: "Analysis of how different factors affect abesenteeism at work"
author: "The absentbees <br> Niamh Wexelstein, Jason Gunn, Samuel Buxton, Xuyan WANG"
institute: "University of Edinburgh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
Absenteeism <- read_csv("Data/Absenteesim At Work.csv")
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "img/confetti.jpg"
)
```

## Why Chose Absenteesim at work

Our team is investigating the different factors which affect people's absence from work. We'll be looking at the variables like bmi, age, social drinker/smoker etc and see how these affect the chance of a person being absent and the frequency of absence. Most of us will be employed in the future so it is interesting to see the different reasons which will affect our work. It is not an overly ambitious topic to work on as this data fits our expectation as there is enough to create our analysis and prove our hypothesis, it also does not have too much data so it won't be complicated to present.


---

# Introduction to our data
First we would like to introduce our data. i.e. the background of the data and its details:
```{r glimpse, out.width = "25%", out.height = "25%", echo = FALSE}
glimpse(Absenteeism)
```


---

# 

- The data comes University of California Irvine machine learning laboratory, the database was created with records of absenteeism at work from July 2007 to July 2010 at a courier company in Brazil.

- There are 21 variables and 740 people are in the dataframe. 

The variables are: 
- ID
- Reason for Absence (Absences attested by the International Code of Diseases (ICD) stratified into 21 categories) Month of Absence, Day of the week & Seasons
- Transportation Expense 
- Distance from Residence to Work
- Service time and age
- Work load Average/day
- Hit target 
- Disciplinary failure 
- Education (Level of)
- Son (The bizzare variable for number of children)
- Social drinker/ smoker
- Pet (number of pets)
- Weight 
- Height
- Body mass index
- Absenteeism time in hours

---

class: inverse, middle, center

# Data Analysis Plan

---

# What We are Going to Do
Our aim is to analyse how different factors affect your likelihood of attending work, and which factor has the most predominant effect.



---
## To Acheive

- First we take an average to some important numeric values, such as absenteeism time, so that we can have a brief understanding about our data, for instance, what the time period may be like. 

- In this analysis, our group is aiming to use linear graph, box plot, and other various graphs in order to analyse the data better.

- In the end our group is aiming for by using the graphs to suggest relationships between different variables. Also may be fiding the correlations.

---

# Layouts

To start with, we may be interested in age linked with absent time taken.

- As age and absenteeism at work may have a casual relationship.

.pull-left[
What we are going to do $^*$
]
.pull-right[
- We can first group absentee time by age and find averages to find ooutliers and then find patterns using graphs
]

.footnote[
[*] We may suggest a few reasons based on the results. Why this is happened.
]

---

# Table and Bar plot

.pull-left[

```{r boring-analysis, out.width = "75%", out.height = "75%", echo = FALSE, fig.align = "centre"}

Absenteeism %>%
  select(Age, Absenteeism_time_in_hours) %>%
  group_by(Age) %>%
  summarise(mean_absent_time_hours = mean(Absenteeism_time_in_hours))
  


```
]

.pull-right[
```{r more-boring-analysis, echo = FALSE, fig.align = "centre"}

Absenteeism %>%
  select(Age, Absenteeism_time_in_hours) %>%
  group_by(Age) %>%
  summarise(mean_absent_time_hours = mean(Absenteeism_time_in_hours)) %>%
  ggplot() +
  geom_col(mapping = aes(x = Age, y = mean_absent_time_hours, fill = Age)) +
  labs(
  title = "Average absenteeism time vs age",
  subtitle = "740 Observations at a brazilian courier company: 07/2007 - 07/2010",
  y = "Average absenteeism time in hours",
  x = "Age"
)
```
]
---

# Hex Plot
```{r even-more-boring-analysis, out.width = "60%", out.height = "60%", echo = FALSE, fig.align = "centre"}
ggplot() +
geom_hex(data = Absenteeism, mapping = aes(x = Absenteeism_time_in_hours , y = Age )) +
labs(
  title = "Absenteeism time vs age",
  subtitle = "740 Observations at a brazilian courier company: 07/2007 - 07/2010",
  y = "Age",
  x = "Absentee time in hours"
)
```

.footnote[
- There are only `r nrow(filter(Absenteeism, Age >= 40))` People above the age of 40 in the workforce out of 740
- And only `r nrow(filter(Absenteeism, Age >= 50))` People above the age of 50 in the workforce
]

---

# Conclusions and Hypothesis
In  conclusion
- Firstly by the table with age and average absenteeism hours, we can clearly see that people aged 34 have the longest absenteeism hours with 16.4 hours and people with age 32 have the least absenteeism hours with 3.69. 

- Secondly

- Our hypothesis



---

# Feeling adventurous?

- Want to find out more about `absenteesim at work`? See https://www.kaggle.com/datasets/saumitgp/absenteeism-at-work.

