# Absenteeism at work
======================
## By we are not here

## Introduction:
We are studying Absenteeism at work, because we believe the majority of us will work in the future, and it will be interesting for us to investigate  different aspects affecting absenteeism.
We got our data set from website 'Kaggle' which they obtained from 'UCI machine learning repository'. The link is following: 'https://archive.ics.uci.edu/ml/datasets/Absenteeism+at+work'. There are 21 variables and 740 observations.

```{r load-data, include=FALSE}
library(tidyverse)
library(tidymodels)
library(knitr)
library(xaringanthemer)
library(pROC)
absenteeism <- read_csv("C:/Users/jasgu/Documents/.Uni Stuff Ig/IDS project/Project-weNotHere/Data/Absenteesim At Work.csv")
```

```{r glimpse data, echo = FALSE}
glimpse(absenteeism)
```

### Cleaning our data:
The first step we took in our project was cleaning the data. We deleted some of the variables: 'Hit target', 'Disciplinary failure', and 'Reason for absence'. We did this because: 'Hit target' is a vague variable name we couldn't make sense of; 'Disciplinary failure' wasn't relevant for our investigation and 'Reason for absence' because we believe it is not important to know why people are absent, only the factors influencing absenteeism. We also find it may be useful to get rid of the people with 0 hours of absenteeism, because we are only interested in people who have been absent.

```{r removing unneeded variables and zero absent hours,}
absenteeism <- select(absenteeism, -c(Reason_for_absence, Hit_target, Disciplinary_failure) ) 
absenteeism <- filter(absenteeism, Absenteeism_time_in_hours != 0)
glimpse(absenteeism)
```

After deleting our variables, we change the name of the column called 'Son' to children as this is more representative. Also we modified the variable 'Day of the Week' as originally Monday was two, Friday was six, so to make more sense, we changed Monday to one and Tuesday to two, etc. 

```{r making more sense of the data}
absenteeism <- filter(absenteeism, Month_of_absence != 0)
absenteeism$Day_of_the_week <- absenteeism$Day_of_the_week -1  # Changing days of the week to 1 = Monday, 5 = Friday
colnames(absenteeism)[colnames(absenteeism) == "Son"] <- "Children"
glimpse(absenteeism)
```

### Splitting our data set:

```{r splitting the dataset}
time_absent <- select(absenteeism,ID, Month_of_absence, Day_of_the_week, Seasons, Absenteeism_time_in_hours)
non_time_absent <- select(absenteeism, -c(Month_of_absence, Day_of_the_week, Seasons))
glimpse(time_absent)
glimpse(non_time_absent)
```


## Time-based absence:


### Renaming our variables:

```{r renaming variables}
time_absent <- time_absent %>%
  mutate(
    Seasons = case_when(
      Seasons == "1" ~ "Summer",
      Seasons == "2" ~ "Autumn",
      Seasons == "3" ~ "Winter",
      Seasons == "4" ~ "Spring"
    ))
time_absent <- time_absent %>%
  mutate(
    Day_of_the_week = case_when(
      Day_of_the_week == "1" ~ "Monday",
      Day_of_the_week == "2" ~ "Tuesday",
      Day_of_the_week == "3" ~ "Wednesday",
      Day_of_the_week == "4" ~ "Thursday",
      Day_of_the_week == "5" ~ "Friday"
  ))
glimpse(time_absent)
```

### Seasons:

```{r seasons}
season_totals <- time_absent %>%
  group_by(Seasons, ID)%>%
  summarise(total_hours_season = sum(Absenteeism_time_in_hours))
glimpse(season_totals)
```

```{r seasons bar chart, echo = FALSE}
season_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Seasons , y = total_hours_season, fill = Seasons)) +
  labs(
    title = "Time spent off work per season",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Season"
) 
  
```

### Months:

```{r months in winter}
month_totals <- time_absent %>%
  filter(Seasons == "Autumn") %>%
  group_by(Month_of_absence) %>%
  summarise(total_hours_month = sum(Absenteeism_time_in_hours))
  glimpse(month_totals)

```

```{r renaming months}
month_totals <- month_totals %>%
  mutate(
    Month_of_absence = case_when(
      Month_of_absence == "1" ~ "January",
      Month_of_absence == "2" ~ "February",
      Month_of_absence == "3" ~ "March",
      Month_of_absence == "12" ~ "December"
    ))
glimpse(month_totals)
```


```{r months bar chart, echo = FALSE}
month_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Month_of_absence , y = total_hours_month, fill = Month_of_absence)) +
  labs(
    title = "Time spent off work per month in winter",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Month"
) 
  
```

### Day of the week

```{r day totals}
day_totals <- time_absent %>%
  group_by(Day_of_the_week) %>%
  summarise(total_hours_day = sum(Absenteeism_time_in_hours))
  glimpse(day_totals)
```

```{r days bar chart, echo = FALSE}
day_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Day_of_the_week , y = total_hours_day, fill = Day_of_the_week)) +
  labs(
    title = "Time spent off work per weekday in winter",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Day of the week"
) 
  
```

## Non time-based absence:

```{r reminder of non time-based}
glimpse(non_time_absent)
```

### Grouping the data set:

```{r grouping by ID, echo = FALSE}
non_time_absent <- non_time_absent %>%
  group_by(ID) %>%
  summarise(
    Transportation_expense = mean(Transportation_expense),
    Distance_from_Residence_to_Work = mean(Distance_from_Residence_to_Work),
    Service_time = mean(Service_time),
    Age = mean(Age),
    Work_load_Average_day = mean(Work_load_Average_day),
    Education = mean(Education),
    Children = mean(Children),
    Social_drinker = mean(Social_drinker),
    Social_smoker = mean(Social_smoker),
    Pet = mean(Pet),
    Weight = mean(Weight),
    Height = mean(Height),
    Body_mass_index = mean(Body_mass_index),
    Absenteeism_time_in_hours = sum(Absenteeism_time_in_hours))
glimpse(non_time_absent)
```

### Age plot:

```{r age plot, echo = FALSE}

non_time_absent %>%
  select(Age, Absenteeism_time_in_hours) %>%
  group_by(Age) %>%
  summarise(mean_absent_time_age = mean(Absenteeism_time_in_hours)) %>%
  ggplot() +
  geom_col(mapping = aes(x = Age, y = mean_absent_time_age, fill = Age)) +
  labs(
  title = "Average time off work vs age",
  subtitle = "Employees at a brazilian courier company: 07/2007 - 07/2010",
  y = "Average time off work in hours",
  x = "Age"
)
```

 There are only `r nrow(filter(non_time_absent, Age >= 40))` People above the age of 40 in the workforce out of 740
 And only `r nrow(filter(non_time_absent, Age >= 50))` People above the age of 50 in the workforce
 
### Education plot :

```{r education box plot, echo = FALSE, warning = FALSE}
ggplot(data = non_time_absent, aes(x = Education, y = Absenteeism_time_in_hours)) +
  geom_boxplot() + 
  facet_wrap(~ Education) +
  ylim(0, 75) +
  labs(title = "Absenteeism time based on education level", 
       x = "Education level",
       y = "absenteeism time (hours)") +
  coord_flip()
```
 
```{r removing education and age}
non_time_absent <- select(non_time_absent, -c(Age, Education, ID))
glimpse(non_time_absent)
```

## Multiple regression model 1:

### Hypothesis:

```{r more complex regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index, y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Title",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "X",
    y = "Predicted time off work (hours)"
  )

lm(formula = Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index , data = non_time_absent) %>%
  summary()
```



## Multiple regression model 2:

### Hypothesis:

```{r simple regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Children + Pet + Social_drinker + Social_smoker, 
    y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Hours taken off work",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "Number of children & pets combined with drinking and smoking (1 = yes) ",
    y = "Predicted time off work (hours)"
  )

simple_regession <- lm(formula = Absenteeism_time_in_hours ~ Children + Social_smoker + Pet + Social_drinker ,
   data = non_time_absent) %>%
  summary()
```

### Smokers box plot

```{r change smoking variables, echo = FALSE, warning = FALSE}
non_time_absent_categorical <- non_time_absent %>%
  mutate(
    Social_smoker = case_when(
      Social_smoker == "1" ~ "Smoker",
      Social_smoker == "0" ~ "Non-smoker" ))
```

```{r smokers box plot, echo = FALSE}
ggplot(data = non_time_absent_categorical, mapping = aes(x = Absenteeism_time_in_hours, y = Social_smoker))+
  geom_boxplot() +
  coord_flip()+
  labs(
    title = "Time to complete 5K and 10 Mile run",
    subtitle = "19,961 runners in the 2019 Cherry Blossom Run",
    y = "Sex",
    x = "Time to complete the Race, seconds",
    colour = "Sex"
  )
```

Smokers make up `r round(((7/33) *100), digits = 1) ` percent of the workforce, so we believe there is enough data to draw a conclusion.

### Residual plots:

```{r complex regression residuals, echo = FALSE}

complex_regression_fit <- linear_reg()%>%
  set_engine("lm") %>%
  fit(Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index , data = non_time_absent)

complex_regression_aug <- augment(complex_regression_fit$fit)

ggplot(complex_regression_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Predicted hours off work", y = "Residuals")


```

```{r simple regression residuals, echo = FALSE}

simple_regression_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit( Absenteeism_time_in_hours ~ Children + Social_smoker + Pet + Social_drinker ,
   data = non_time_absent)


simple_regression_aug <- augment(simple_regression_fit$fit)

ggplot(simple_regression_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Predicted hours off work", y = "Residuals")
```

Slight less fan pattern (though slight fan pattern lmao - just say it's an improvement over the other one)

## Conclusion:

Most of it is really random lol

Thursdays in December and also smoke apparently lol

That is all Im going to sleep now :))


