---
title: "Absenteeism-at-work"
subtitle: "Analysis of how different factors affect absenteeism at work"
author: "The absentbees <br> Niamh Wexelstein, Jason Gunn, Samuel Buxton, Xuyan Wang"
institute: "University of Edinburgh"
date: "`r Sys.Date()`" 
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    seal: false 
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---
```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
absenteeism <- read_csv("/Users/samuelbuxton540/Desktop/uni?/ids/r studio commits/Project-weNotHere/Data/Absenteesim At Work.csv")
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "presentation.title.jpeg"
)
```

class: title-slide, top, left, inverse

.pull-left[

]

.pull-right[
# **`r rmarkdown::metadata$title`**
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$institute`
### `r Sys.Date()`
]

---

## What are we going to do?

We are analysing the different factors that affect people being absent from work. We thought this would be interesting to look at because the majority of us will get employed one day and we wish to know the 
different factors that are likely to affect our enthusiasm to turn up to work.


---

# Hypothesis

We are going to analyse 3 variables in particular. We think that the further you are from work, the more children you have, and the lower your education makes you more likely to be absent from work.


---
#Our dataset

```{r glimpse, out.width = "20%", out.height = "20%", echo = FALSE}
head(absenteeism, 10)
```

---
# Cleaning the dataset

```{r Cleaning, eval = TRUE }

# Getting rid of unneeded/unwanted variables 
absenteeism <- select(absenteeism, -c(Reason_for_absence, Hit_target, Disciplinary_failure) ) 
absenteeism <- filter(absenteeism, Month_of_absence != 0)
absenteeism$Day_of_the_week <- absenteeism$Day_of_the_week -1  # Changing days of the week to 1 = Monday, 5 = Friday
colnames(absenteeism)[colnames(absenteeism) == "Son"] <- "Children"
absenteeism
```

---

#Methods

* We removed the variables "reason for absence", "hit target" and "disciplinary failure" as they aren't relevant to the purpose of our report
* Box plot of some data such as "age" and "absenteesim time in hours"
* Finding averages of some interesting variables by statistical method
* Logistic regression model to find a linear relationship between different variables. (set Absenteesim hours as y for our models)
---
class: middle, inverse

#Introductory plots

---
##Average absenteeism split by different variables
```{r warning = FALSE, out.width = "70%", out.height = "70%", echo = FALSE, fig.align = "centre"}
ggplot(data = absenteeism, aes(x = Education, y = Absenteeism_time_in_hours)) +
  geom_boxplot() + 
  facet_wrap(~ Education) +
  ylim(0, 75) +
  labs(title = "Absenteeism time based on education level", 
       x = "Education level",
       y = "absenteeism time (hours)") +
  coord_flip() 


```
---
#Observations
The graph in the previous slide shows that the lower your education level, the less likely you are to be absent. Our conjecture is that this is because people with lower education levels have fewer job opportunities, so therefore feel the need to work harder to maintain the job they have.
---
class: middle, inverse

#Grouping
---
#Grouping social factors
```{r social factors}
absenteeism_social <- absenteeism %>%
  select(Education, Children, Social_drinker, Social_smoker, Pet, Absenteeism_time_in_hours)
absenteeism_social
```

---
#Grouping biological factors
```{r biological factors}
absenteeism_biological<- absenteeism %>%
  select(Age, Weight, Height, Body_mass_index, Absenteeism_time_in_hours)
absenteeism_biological
```

---
#Grouping time factors
```{r time-based factors}
absenteeism_time <- absenteeism %>%
  select(Month_of_absence, Day_of_the_week, Seasons, Service_time, Absenteeism_time_in_hours)
absenteeism_time
```

---
#Grouping geographical factors
```{r geographical factors}
absenteeism_geog <- absenteeism %>%
  select(Transportation_expense, Distance_from_Residence_to_Work, Absenteeism_time_in_hours)
absenteeism_geog
```

---
class: middle, inverse

#Further plots/analysis
---
#Geographical linear regression
```{r geographical linear regression}
geog_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work, data = absenteeism_geog) 
geog_fit
```

---
#Average absenteeism vs distance from residence to work
```{r}
ggplot(data = absenteeism_geog, aes(x =  Distance_from_Residence_to_Work, y = Absenteeism_time_in_hours)) + 
  geom_hex() 

geog_fit_aug <- augment(geog_fit$fit)
```

---
#Residual plot
```{r}
ggplot(geog_fit_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "gray", lty = "dashed") +
  labs(x = "Predicted height", y = "Residuals")
```

---
#Observations/Analysis
The linear regression shows that for each unit increase in distance from residence to work, you are roughly 3% less likely to be absent from work & for each unit increase in transportation expense, you are 42% more likely to be absent from work. This assertion is backed up by the data in the hex plot, and by the random distribution of points in the residual plot.
---

#Evalutions
- The main flaw with the investigation is the use of only one data set. This makes the results less reliable as this one company may show results that aren't representative of an average workforce. We could improve this investigation through the use of two (or more) data sets.
---

#Conclusions
In conclusion, in order to lower your likelihood of being absent from work, it is best to live as far away from work and lower your transportation expenses as much as possible.
```


