---
title: "Absenteeism-at-work"
subtitle: "Analysis of how different factors affect absenteeism at work"
author: "Niamh Wexelstein, Jason Gunn, Samuel Buxton, Xuyan Wang"
institute: "University of Edinburgh"
date: "`r Sys.Date()`" 
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    seal: false 
    nature:
      ratio: "32:18"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
absenteeism <- read_csv("C:/Users/jasgu/Documents/.Uni Stuff Ig/IDS project/Project-weNotHere/Data/Absenteesim At Work.csv")
```

```{r include=FALSE}
style_xaringan(
  title_slide_background_image = "presentation.title.jpeg"
)
```

class: title-slide, top, left, inverse

.pull-left[

]

.pull-right[
# **`r rmarkdown::metadata$title`**
## **`r rmarkdown::metadata$subtitle`**
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$institute`
### `r Sys.Date()`
]

---

## What are we going to do?

We are analysing the different factors that affect people being absent from work. We thought this would be interesting to look at because the majority of us will get employed one day and we wish to know the different factors that are likely to affect our enthusiasm to turn up to work.


---

# Hypothesis

We are going to analyse 3 variables in particular. We think that:
- The further you are from work, the less likely you are to attend.
- The more children you have, the more likely you are to be absent.
- The lower your education, the more likely you are to be absent from work.


---

#Our dataset

```{r glimpse, out.width = "20%", out.height = "20%", echo = FALSE}
head(absenteeism, 10)
```
---
class: middle, inverse

# Cleaning the dataset
---
<<<<<<< HEAD
=======

# We removed unwanted variables


# We removed unwanted variables

```{r Cleaning, eval = TRUE, echo = FALSE }

absenteeism <- select(absenteeism, -c(Reason_for_absence, Hit_target, Disciplinary_failure) ) 
absenteeism <- filter(absenteeism, Month_of_absence != 0)
absenteeism$Day_of_the_week <- absenteeism$Day_of_the_week -1  # Changing days of the week to 1 = Monday, 5 = Friday
colnames(absenteeism)[colnames(absenteeism) == "Son"] <- "Children"
absenteeism
```

---
<<<<<<< HEAD

# We split the dataset
```{r splitting the dataset, echo = FALSE, out.extra = 2000}
=======
# We split the dataset
```{r splitting the dataset}
>>>>>>> c2a4f969570ac2b2894002a8fa70ec79586c7e09
time_absent <- select(absenteeism,ID, Month_of_absence, Day_of_the_week, Seasons, Absenteeism_time_in_hours)
non_time_absent <- select(absenteeism, -c(Month_of_absence, Day_of_the_week, Seasons))
glimpse(time_absent)
glimpse(non_time_absent)
```
<<<<<<< HEAD
=======
---

# We renamed variables
```{r renaming variables}
time_absent <- time_absent %>%
  mutate(
    Seasons = case_when(
      Seasons == "1" ~ "Summer",
      Seasons == "2" ~ "Autumn",
      Seasons == "3" ~ "Winter",
      Seasons == "4" ~ "Spring"
    ))
time_absent <- time_absent %>%
  mutate(
    Day_of_the_week = case_when(
      Day_of_the_week == "1" ~ "Monday",
      Day_of_the_week == "2" ~ "Tuesday",
      Day_of_the_week == "3" ~ "Wednesday",
      Day_of_the_week == "4" ~ "Thursday",
      Day_of_the_week == "5" ~ "Friday"
  ))
glimpse(time_absent)
```
--- 


# Seasons:

```{r seasons}
season_totals <- time_absent %>%
  group_by(Seasons, ID) %>%
  summarise(total_hours_season = sum(Absenteeism_time_in_hours))
glimpse(season_totals)
```
---


# We renamed variables
```{r renaming variables, echo = FALSE}
time_absent <- time_absent %>%
  mutate(
    Seasons = case_when(
      Seasons == "1" ~ "Summer",
      Seasons == "2" ~ "Autumn",
      Seasons == "3" ~ "Winter",
      Seasons == "4" ~ "Spring"
    ))
time_absent <- time_absent %>%
  mutate(
    Day_of_the_week = case_when(
      Day_of_the_week == "1" ~ "Monday",
      Day_of_the_week == "2" ~ "Tuesday",
      Day_of_the_week == "3" ~ "Wednesday",
      Day_of_the_week == "4" ~ "Thursday",
      Day_of_the_week == "5" ~ "Friday"
  ))
glimpse(time_absent)
```

---

# Seasons:

```{r seasons}
season_totals <- time_absent %>%
  group_by(Seasons, ID) %>%
  summarise(total_hours_season = sum(Absenteeism_time_in_hours))
glimpse(season_totals)
```

---

# Time spent off work per season
```{r seasons bar chart, echo = FALSE}
season_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Seasons , y = total_hours_season, fill = Seasons)) +
  labs(
    title = "Time spent off work per season",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Season"
) 
  
```

---

# Months


Since Autumn has the least absent hours we looked at the months in Autumn:

```{r months in autumn}
month_totals <- time_absent %>%
  filter(Seasons == "Autumn") %>%
  group_by(Month_of_absence) %>%
  summarise(total_hours_month = sum(Absenteeism_time_in_hours))
  glimpse(month_totals)

```


---

# Renaming months



```{r renaming months}
month_totals <- month_totals %>%
  mutate(
    Month_of_absence = case_when(
      Month_of_absence == "1" ~ "January",
      Month_of_absence == "2" ~ "February",
      Month_of_absence == "3" ~ "March",
      Month_of_absence == "12" ~ "December"
    ))
glimpse(month_totals)
```


---

# Time spent off work per month in autumn

```{r months bar chart, echo = FALSE}
month_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Month_of_absence , y = total_hours_month, fill = Month_of_absence)) +
  labs(
<<<<<<< HEAD
    title = "Time spent off work per month in autumn",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Month",
    fill = "Month"
) 
  
```


# Day of the week

```{r day totals}
day_totals <- time_absent %>%
  group_by(Day_of_the_week) %>%
  summarise(total_hours_day = sum(Absenteeism_time_in_hours))
  glimpse(day_totals)
```


---

# Time spent off work per weekday in winter
```{r days bar chart, echo = FALSE}
day_totals %>%
ggplot() +
  geom_col(mapping = aes(x = Day_of_the_week , y = total_hours_day, fill = Day_of_the_week)) +
  labs(

    title = "Time spent off work per weekday",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    y = "Hours off work",
    x = "Day of the week",
    fill = "Day of the week"
```

## Non time-based absence:

```{r reminder of non time-based}
glimpse(non_time_absent)
```

---

### Grouping the data set:

```{r grouping by ID, echo = FALSE}
non_time_absent <- non_time_absent %>%
  group_by(ID) %>%
  summarise(
    Transportation_expense = mean(Transportation_expense),
    Distance_from_Residence_to_Work = mean(Distance_from_Residence_to_Work),
    Service_time = mean(Service_time),
    Age = mean(Age),
    Work_load_Average_day = mean(Work_load_Average_day),
    Education = mean(Education),
    Children = mean(Children),
    Social_drinker = mean(Social_drinker),
    Social_smoker = mean(Social_smoker),
    Pet = mean(Pet),
    Weight = mean(Weight),
    Height = mean(Height),
    Body_mass_index = mean(Body_mass_index),
    Absenteeism_time_in_hours = sum(Absenteeism_time_in_hours))
glimpse(non_time_absent)
```

---

# Age plot:

```{r age plot, echo = FALSE}

non_time_absent %>%
  select(Age, Absenteeism_time_in_hours) %>%
  group_by(Age) %>%
  summarise(mean_absent_time_age = mean(Absenteeism_time_in_hours)) %>%
  ggplot() +
  geom_col(mapping = aes(x = Age, y = mean_absent_time_age, fill = Age)) +
  labs(
  title = "Average time off work vs age",
  subtitle = "Employees at a brazilian courier company: 07/2007 - 07/2010",
  y = "Average time off work in hours",
  x = "Age"
)
```
<<<<<<< HEAD

=======
---
There are only `r nrow(filter(non_time_absent, Age >= 40))` People above the age of 40 in the workforce out of 740
 And only `r nrow(filter(non_time_absent, Age >= 50))` People above the age of 50 in the workforce
>>>>>>> c2a4f969570ac2b2894002a8fa70ec79586c7e09
---

# Education plot :

```{r education box plot, echo = FALSE, warning = FALSE}
ggplot(data = non_time_absent, aes(x = Education, y = Absenteeism_time_in_hours)) +
  geom_boxplot() + 
  facet_wrap(~ Education) +
  ylim(0, 75) +
  labs(title = "Absenteeism time based on education level", 
       x = "Education level",
       y = "absenteeism time (hours)") +
  coord_flip()
<<<<<<< HEAD
=======
```

```{r removing education and age}
non_time_absent <- select(non_time_absent, -c(Age, Education, ID))
glimpse(non_time_absent)
```
---
class: middle, inverse

# Regression models
---

## Multiple regression model 1:
```{r more complex regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index, y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "X",
    y = "Predicted time off work (hours)"
  )

lm(formula = Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index , data = non_time_absent) %>%
  summary()
```
--- 


```{r removing education and age}
non_time_absent <- select(non_time_absent, -c(Age, Education, ID))
glimpse(non_time_absent)

## Multiple regression model 2:
```{r simple regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Children + Pet + Social_drinker + Social_smoker, 
    y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Hours taken off work",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "Number of children & pets combined with drinking and smoking (1 = yes) ",
    y = "Predicted time off work (hours)"
  )

simple_regession <- lm(formula = Absenteeism_time_in_hours ~ Children + Social_smoker + Pet + Social_drinker ,
   data = non_time_absent) %>%
  summary()
```
---

### Smokers box plot
```{r change smoking variables, echo = FALSE, warning = FALSE}
non_time_absent_categorical <- non_time_absent %>%
  mutate(
    Social_smoker = case_when(
      Social_smoker == "1" ~ "Smoker",
      Social_smoker == "0" ~ "Non-smoker" ))
```
---

```{r smokers box plot, echo = FALSE}
ggplot(data = non_time_absent_categorical, mapping = aes(x = Absenteeism_time_in_hours, y = Social_smoker))+
  geom_boxplot() +
  coord_flip()+
  labs(
    title = "Time to complete 5K and 10 Mile run",
    subtitle = "19,961 runners in the 2019 Cherry Blossom Run",
    y = "Sex",
    x = "Time to complete the Race, seconds",
    colour = "Sex"
  )

```
---

class: middle, inverse

# Regression models
---

## Multiple regression model 1:

```{r more complex regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index, y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "X",
    y = "Predicted time off work (hours)"
  )
```

---

## R sqared for regression model 1

```{r AAAAAA, echo = FALSE, out.extra = 4000}
lm(formula = Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index , data = non_time_absent) %>%
  summary()
```

---

## Multiple regression model 2:

```{r simple regression model, echo = FALSE}

ggplot(data = non_time_absent, aes(x = Children + Pet + Social_drinker + Social_smoker, 
    y = Absenteeism_time_in_hours )) +
  geom_point(position = "jitter") +
  geom_smooth(method = "lm", se = FALSE) + 
  labs(
    title = "Hours taken off work",
    subtitle = "Employees of a Brazilian courier company July 2007-10",
    x = "Number of children & pets combined with drinking and smoking (1 = yes) ",
    y = "Predicted time off work (hours)"
  )
```

---

## R squared for regression model 2
```{r EEEEEE, echo = FALSE, out.extra = 2000}
lm(formula = Absenteeism_time_in_hours ~ Children + Social_smoker + Pet + Social_drinker ,
   data = non_time_absent) %>%
  summary()
```

---

### Smokers box plot
```{r change smoking variables, echo = FALSE, warning = FALSE}
non_time_absent_categorical <- non_time_absent %>%
  mutate(
    Social_smoker = case_when(
      Social_smoker == "1" ~ "Smoker",
      Social_smoker == "0" ~ "Non-smoker" ))
```


```{r smokers box plot, echo = FALSE}
ggplot(data = non_time_absent_categorical, mapping = aes(x = Absenteeism_time_in_hours, y = Social_smoker))+
  geom_boxplot() +
  coord_flip()+
  labs(
    title = "Smoking status vs Hours spent off work",
    subtitle = "Employees of a Brazilian courier company 07/07 - 07/10",
    x = "Hours spent off work",
    y = "Smoking status",
  )
```

Smokers make up `r round(((7/33) *100), digits = 1) ` percent of the workforce, so we believe there is enough data to draw a conclusion.
---
class: middle, inverse

# Regression models
---

# Complex regression residual
=======
Smokers make up `r round(((7/33) *100), digits = 1) ` percent of the workforce, so we believe there is enough data to draw a conclusion.
---

### Residual plots:

```{r complex regression residuals, echo = FALSE}

complex_regression_fit <- linear_reg()%>%
  set_engine("lm") %>%
  fit(Absenteeism_time_in_hours ~ Transportation_expense + Distance_from_Residence_to_Work + Children
   + Service_time + Work_load_Average_day + Social_drinker + Social_smoker + Pet
   + Weight + Height + Body_mass_index , data = non_time_absent)

complex_regression_aug <- augment(complex_regression_fit$fit)

ggplot(complex_regression_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Predicted hours off work", y = "Residuals")


```
---


# Simple model residual

=======

```{r simple regression residuals, echo = FALSE}

simple_regression_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit( Absenteeism_time_in_hours ~ Children + Social_smoker + Pet + Social_drinker ,
   data = non_time_absent)


simple_regression_aug <- augment(simple_regression_fit$fit)

ggplot(simple_regression_aug, mapping = aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", lty = "dashed") +
  labs(x = "Predicted hours off work", y = "Residuals")
```
---

=======
#Evaluations
- The main flaw with the investigation is the use of only one data set. This makes the results less reliable as this one company may show results that aren't representative of an average workforce. We could improve this investigation through the use of two (or more) data sets
- It must also be taken into account that this data set was made 13 years ago (pre-covid), so a more up-to-date data set could provide more timely results

## Conclusion
- Most of the data it is really random
- The more you smoke, the less you drink, the less children and more pets you have, the less likely you are to be absent from work
- Statistically, Thursdays in December have the lowest rates of absence
--- 